Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-09-02 paper/zhang-cook-laa-langrene-menendez.Rmd
Added scagnostics reference.

Let $\mathbf{X}_{n \times p}$ be the data matrix with $n$ observations in $p$ dimensions. A $d$-dimensional projection is a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$ defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection matrix. We define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ onto a scalar value. This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection. An interesting projection shows structures that are non-normal since theoretical proofs from @diaconis1984asymptotics have shown that projections tend to be normal, as $n$ and $p$ approach infinity under certain conditions. There have been many index functions proposed in the literature, here are a few examples: early indexes that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; LDA index [@lee2005projection] and PDA [@lee2010projection] index for supervised classification; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indexes [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.

>>>>

Let $\mathbf{X}_{n \times p}$ be the data matrix with $n$ observations in $p$ dimensions. A $d$-dimensional projection is a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$ defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection matrix. We define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ onto a scalar value. This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection. An interesting projection shows structures that are non-normal since theoretical proofs from @diaconis1984asymptotics have shown that projections tend to be normal, as $n$ and $p$ approach infinity under certain conditions. There have been many index functions proposed in the literature, here are a few examples: early indexes that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; LDA index [@lee2005projection] and PDA [@lee2010projection] index for supervised classification; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indexes [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures [@scag,@WW08].

### 2021-09-02 paper/zhang-cook-laa-langrene-menendez.tex
Added scagnostics reference.

Let \(\mathbf{X}_{n \times p}\) be the data matrix with \(n\)
observations in \(p\) dimensions. A \(d\)-dimensional projection is a
linear transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\)
defined as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where
\(\mathbf{Y}_{n \times d}\) is the projected data and
\(\mathbf{A}_{p\times d}\) is the projection matrix. We define
\(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index
function that maps the projected data \(\mathbf{Y}\) onto a scalar
value. This is commonly known as the projection pursuit index function,
or just index function, and is used to measure the ``interestingness''
of a given projection. An interesting projection shows structures that
are non-normal since theoretical proofs from
\citet{diaconis1984asymptotics} have shown that projections tend to be
normal, as \(n\) and \(p\) approach infinity under certain conditions.
There have been many index functions proposed in the literature, here
are a few examples: early indexes that can be categorised as measuring
the \(L^2\) distance between the projection and a normal distribution:
Legendre index \citep{friedman1974projection}, Hermite index
\citep{hall1989polynomial}, and natural Hermite index
\citep{cook1993projection}; chi-square index \citep{posse1995projection}
for detecting spiral structure; LDA index \citep{lee2005projection} and
PDA \citep{lee2010projection} index for supervised classification;
kurtosis index \citep{Loperfido2020} and skewness index
\citep{Loperfido2018} for detecting outliers in financial time series;
and most recently, scagnostic indexes \citep{laa2020using} for
summarising structures in scatterplot matrices based on eight scagnostic
measures.

@@

{\centering \includegraphics[width=0.7\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/tour-path}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-search-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-interp-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-pca}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-pca-animated-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-tour-1}

@@

As mentioned previously, the original \(p \times d\) dimension space can
be simulated via randomly generated bases in the \pkg{geozoo}
\citep{geozoo} package. While the PCA plot projects the bases from the
direction that maximises the variance, the tour plot displays the
original high-dimensional space from various directions using animation.
Figure \ref{fig:toy-tour} shows some frames from the tour plot of the
same two optimisations in its original 5D space.

@@

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-torus-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/interruption-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/polish-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/noisy-better-geo-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/kol-result-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/flip-sign-1}

@@

\address{%
H.Sherry Zhang\\
Monash University\\%
Department of Econometrics and Business Statistics\\
%
Melbourne, Australia\\%
%
\href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}%
}

\address{%
Dianne Cook\\
Monash University\\%
Department of Econometrics and Business Statistics\\
%
Melbourne, Australia\\%
%
\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}%
}

\address{%
Ursula Laa\\
University of Natural Resources and Life Sciences\\%
Institute of Statistics\\
%
Vienna, Austria\\%
%
\href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}%
}

\address{%
Nicolas Langrené\\
CSIRO Data61\\%
34 Village Street, Docklands VIC 3008 Australia\\
%
Melbourne, Australia\\%
%
\href{mailto:nicolas.langrene@csiro.au}{\nolinkurl{nicolas.langrene@csiro.au}}%
}

\address{%
Patricia Menéndez\\
Monash University\\%
Department of Econometrics and Business Statistics\\
%
Melbourne, Australia\\%
%
\href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}%
}

>>>>

Let \(\mathbf{X}_{n \times p}\) be the data matrix with \(n\)
observations in \(p\) dimensions. A \(d\)-dimensional projection is a
linear transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\)
defined as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where
\(\mathbf{Y}_{n \times d}\) is the projected data and
\(\mathbf{A}_{p\times d}\) is the projection matrix. We define
\(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index
function that maps the projected data \(\mathbf{Y}\) onto a scalar
value. This is commonly known as the projection pursuit index function,
or just index function, and is used to measure the ``interestingness''
of a given projection. An interesting projection shows structures that
are non-normal since theoretical proofs from
\citet{diaconis1984asymptotics} have shown that projections tend to be
normal, as \(n\) and \(p\) approach infinity under certain conditions.
There have been many index functions proposed in the literature, here
are a few examples: early indexes that can be categorised as measuring
the \(L^2\) distance between the projection and a normal distribution:
Legendre index \citep{friedman1974projection}, Hermite index
\citep{hall1989polynomial}, and natural Hermite index
\citep{cook1993projection}; chi-square index \citep{posse1995projection}
for detecting spiral structure; LDA index \citep{lee2005projection} and
PDA \citep{lee2010projection} index for supervised classification;
kurtosis index \citep{Loperfido2020} and skewness index
\citep{Loperfido2018} for detecting outliers in financial time series;
and most recently, scagnostic indexes \citep{laa2020using} for
summarising structures in scatterplot matrices based on eight scagnostic
measures \citep[\citet{WW08}]{scag}.

@@

{\centering \includegraphics[width=0.7\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/tour-path}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-search-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-interp-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-pca}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-pca-animated-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-tour-1}

@@

As mentioned previously, the original \(p \times d\) dimension space can
be simulated via randomly generated bases in the \pkg{geozoo}
\citep{geozoo} package. While the PCA plot projects the bases from the
direction that maximises the variance, the tour plot displays the
original high-dimensional space from various directions using animation.
Figure \ref{fig:toy-tour} shows some frames from the tour plot of the
same two optimisations in its original space.

@@

{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-torus-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/interruption-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/polish-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/noisy-better-geo-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/kol-result-1}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/flip-sign-1}

@@

\address{%
H.Sherry Zhang\\
Monash University\\%
Department of Econometrics and Business Statistics\\
%
Melbourne, Australia%
%
\\\href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}
}

\address{%
Dianne Cook\\
Monash University\\%
Department of Econometrics and Business Statistics\\
%
Melbourne, Australia%
%
\\\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}
}

\address{%
Ursula Laa\\
University of Natural Resources and Life Sciences\\%
Institute of Statistics\\
%
Vienna, Austria%
%
\\\href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}
}

\address{%
Nicolas Langrené\\
CSIRO Data61\\%
34 Village Street, Docklands VIC 3008 Australia\\
%
Melbourne, Australia%
%
\\\href{mailto:nicolas.langrene@csiro.au}{\nolinkurl{nicolas.langrene@csiro.au}}
}

\address{%
Patricia Menéndez\\
Monash University\\%
Department of Econometrics and Business Statistics\\
%
Melbourne, Australia%
%
\\\href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}
}
